---
title: "Data Science and Analytics 2022"
author: "İrem Uslu - 170709043"
output:
  html_document: default
  pdf_document: default
---

<style type="text/css">
  body{
  font-size: 11pt;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!--IMPORTANT!!!!! Please fill out the sections that say:"write your code here"-->
<!--6. For every step above, write down one or two sentences description-->

<br>

```{r echo=T, results='hide', warning=FALSE, message=FALSE}
  library(ggplot2)
  library(tidyverse)
  library(dplyr)
  library(plyr)
```

<br>

### 1) Download the dataset from the link below and import the dataset to R as a dataframe named ***'alcohol_data'***.

<br>

***https://www.kaggle.com/datasets/whenamancodes/alcohol-effects-on-study***

<br>
```{r}
    alcohol_data <- read.csv("Portuguese.csv" , header = TRUE , sep = ",")
```

<br>

### 2)

#### A) Create a new dataframe as ***'alcohol_data_small'*** from ***'alcohol_data'*** by choosing at least 5 columns (at least one factor and one numeric). 

<br>

```{r}
  alcohol_data_small <- alcohol_data [, c(1:4,6,13:14,16:17,27:28,31:33)]
```


<br>

#### B) Check and ***re-format*** column/ variable types (factor, charecter, date, integer etc).

<br>

```{r}
  str(alcohol_data_small)
```

The dataframe I created has 14 observations.These 14 observations consist of integer and character data types.


```{r}
  alcohol_data_small <- transform(alcohol_data_small ,
                                  traveltime = as.factor(traveltime),
                                  studytime = as.factor(studytime),
                                  Walc = as.factor(Walc)
                                  )
```

```{r} 
  str(alcohol_data_small)

```

I converted travel time, daily alcohol consumption and study time into factor data type.

<br>

#### C) By using the given dictionary in kaggle, rename the variables as a more detailed version (e.g. 'F' => 'Female').

<br>

```{r}
  alcohol_data_small <- transform(alcohol_data_small ,
                                  school = as.factor(mapvalues(school , c("GP","MS"),c("Gabriel Pereira",
                                                                                       "Mousinho da Silveira"))),
                                  sex = as.factor(mapvalues(sex , c("F","M") , c("Female" , "Men"))),
                                  address = as.factor(mapvalues(address , c("R" , "U") , c("Rural" , "Urban"))),
                                  Pstatus = as.factor(mapvalues(Pstatus , c("A","T") , c("Together","Apart"))),
                                  Dalc = as.factor(mapvalues(Dalc , c(1,2,3,4,5) , c("Very Low","Low","Medium","High","Very High"))))
                                  
```

```{r}
summary(alcohol_data_small)
```

I used more understandable names for some variables.I also examined how many observations I have with the summary() function and the data types of the variables.
<br>

### 3) 

#### A) Use ***filter***, ***arrange***, ***select*** to manage your data.
<br>

```{r}
  df <- alcohol_data_small %>% 
                          filter(Pstatus == "Apart" & famsup == "yes" & G3>18)%>%
                          select(school,Pstatus,famsup, G3)%>%
                          arrange(G3)
  df

```


I wanted to list the people whose parents live separately, who received educational support from their families other than school, and who had the highest final score. As a result, I observed that only 1 person met these criteria.


```{r}
  df1 <- alcohol_data_small %>% 
                          filter(Dalc == "Very High" & Walc==5)%>%
                          select(sex,studytime,Dalc,Walc)
  
head(df1,6)

```

I observed both the gender and weekly study duration of students with very high weekly and daily alcohol rates.


```{r}
  df2 <- alcohol_data_small %>% 
                    filter(G1==G2 & G2==G3) %>%
                    select(school,G1,G2,G3)%>%
                    arrange(desc(G3))

  head(df2,6)
  
```

I observed which school the students whose 3 grades were equal to each other were studying in.



<br>

#### B) Use ***mutate*** or ***transmute*** to create a new dataframe and save it as ***'new_data.csv'*** in your current working directory.

<br>

```{r}
  new_df <- alcohol_data_small%>%
                              mutate(newG1 = G1+mean(G1),
                                     newG2 = G2+mean(G2),
                                     newG3 = G3+mean(G3))
  head(new_df[c(15:17)])
```

Using mutate() I got the new variables by adding the average of each score on top of the existing scores, and with these new variables I created a new dataframe called new_df.

```{r}
  write.csv(x = new_df, file = "new_data.csv", row.names = FALSE)
```

I wrote and saved the new dataframe I created to a new csv file with write.csv().

<br>

### 4) Use ***'alcohol_data_small'*** dataframe and ggplot2 library to draw the following plots.

<br>

#### A) Draw a ***histogram*** and a ***bar chart*** for any variables and comment about shape and frequency. 

<br>

```{r}
  ggplot(alcohol_data_small , mapping = aes(x = address, fill = traveltime)) + 
  geom_bar(position = "identity",color="black") +
  scale_fill_brewer(palette = "Pastel2")+
  labs( x = "Student's home address type",
        y = "Count",
        title = "Travel times to school according to the student's home address type")

  
```

Students living in rural areas take longer to reach school than students living in urban areas.
Also, there are no students in rural areas who can reach it in 15 minutes or less.

```{r}

  ggplot(alcohol_data_small, aes(age,fill=school)) + 
  geom_histogram(position="identity",binwidth=0.9,color="pink") +
  scale_fill_brewer(palette = "Pastel2")+
  ggtitle("Histogram graph of the age distribution of students in both schools") +
  xlab("Student's Age") + 
  ylab("Count")

  
```

Most students at the school are between the ages of 16 and 18.Very few students go to school after the age of 20.



```{r}
  ggplot(alcohol_data_small, aes(G3, fill=Dalc)) + 
  geom_histogram(position="identity", binwidth=5,color="black") +
  scale_fill_brewer(palette = "Pastel2")+
  facet_grid(Dalc ~ .) + 
  ggtitle("Final scores according to students' daily alcohol consumption") +
  xlab("Final Score") + 
  ylab("Count")

  
```

I examined the final scores of the students according to their daily alcohol consumption. At the 5th level, that is, the final scores of those who consume alcohol at a very high level decrease as much.

<br>

#### B) Draw a ***boxplot*** and comment about the distribution and outliers. 

<br>

```{r ,warning=FALSE}
  ggplot(alcohol_data_small, mapping = aes(x = studytime, y = G1 , fill=studytime)) + 
  geom_boxplot(outlier.size = 2 ,outlier.shape = 21)+
  stat_summary(fun.y = "mean",shape=4,size=0.2)+
  labs(title = "BoxPlot of First Period Scores by Studying Hours", x="Weekly Studying Times", y="First Period (G1) Scores")+
  scale_fill_brewer(palette = "Pastel2")
```
<br>

I have obtained a box chart where we can comment on the scores of the students in the first exam according to their weekly study time.

While the 4th value, that is, those who work more than 10 hours a week, got a higher score in the exam, the 1st value, that is, those who work less than 2 hours per week, gathered around the lower score.

In those with the 2nd value, that is, those who work between 2 and 5 hours a week, there are values ​​that are much higher and lower than the average. These are outlier values.

In addition, those with the most weekly working hours have the highest average score.

<br>