---
title: "CENG 4515 DATA SCIENCE AND ANALYTICS-HW2"
author: "İrem Uslu"
date: "2022-11-17"
output: html_document


---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Importing libraries

```{r echo=T, results='hide', warning=FALSE, message=FALSE}
  library(ggplot2)
  library(tidyverse)
  library(dplyr)
  library(plyr)
  library(caret)
  library(corrplot)
  library(ROCR)
  library(Rcpp)
  library(gmodels)
  library(pROC)
  library(InformationValue)
  library(factoextra)
  library(car)

```


<br>

### Dowload the "mole_dataset.csv" and data preprocessing.

<br>
```{r}
mole <- read.csv("mole_dataset.csv" , header=TRUE , sep=",")

mole$sex <- NULL
mole$X <- NULL
mole$image_id <- NULL


mole$benign_malignant <-  factor(mole$benign_malignant) 
summary(mole)
```

<br>

### Correlation matrix and heatmap between independent variables.

<br>
```{r}

mole <- na.omit(mole)
cor <- cor(mole[,1:10])
corrplot(cor,method = "circle")

```


<br>
Dark blue dots have strong positive correlations.
For example, it can be said that there is a strong relationship between the x and y pixels and the full size of the image.
Or it can be said that the x and y pixels have a weak positive relationship with the vertices.
<br>

```{r}
heatmap(cor(mole[,-11]))

```

<br>It can be said that there is an increasing relationship from the corners variable to the green variable.<br>




### Build a logistic regression model to raw data first.Train and test dataset are created.



```{r}
#Malignant and benign numbers were checked for logistic regression. 
dataMalignant <- mole %>%
                filter(benign_malignant=="malignant")
                nrow(dataMalignant)

dataBenign <- mole %>% 
              filter(benign_malignant=="benign")
              nrow(dataBenign)


set.seed(100)
dataMalignantIndex <- sample(1:nrow(dataMalignant) , size= 0.70*nrow(dataMalignant))

dataBenignIndex <- sample(1:nrow(dataBenign) , size= 0.70*nrow(dataMalignant))



# Train dataset was created.
trainMalignant <- dataMalignant[dataMalignantIndex,]
trainBenign <- dataBenign[dataBenignIndex,]

trainSet <- rbind(trainMalignant , trainBenign)
table(trainSet$benign_malignant) #Malignant and benign values were equalized.

# Test dataset was created.
testMalignant <- dataMalignant[-dataMalignantIndex,]
testBenign <- dataBenign[-dataBenignIndex,]

testSet <- rbind(testMalignant , testBenign)
table(testSet$benign_malignant) 


```


```{r}
model <- glm(benign_malignant~.,family="binomial", data=trainSet)
summary(model)

```
Variables such as age_approx, blue, green, non_overlapping_rate have a significant effect on determining the type of mole. However, pixel_y,red and full_img_size have no significant effect.


### Confusion Matrix

```{r}
prediction <- predict(model , testSet ,type="response")

cm <- confusionMatrix(testSet$benign_malignant , prediction)
cm

accur <- (cm[1,1] + cm[2,2])/ sum(cm)
accur
```
43 benign and 39 malignant moles were predicted correctly.The accuracy rate of the model I created is 0.67. Correct predictions were made in the test data set at a rate of 0.67.

### Roc Curve

```{r}
rocModel <- roc(testSet$benign_malignant ~ prediction) #control = negative class , case = positive class
```
```{r}
plot(rocModel)

rocModel
```

The larger the area under the ROC curve, the better our model will be, so the higher the prediction power.The AUC value of the model I created is 0.71, so the estimation performance is quite high.



### Apply PCA

```{r }

molePr <- prcomp(mole[,-11] , scale=TRUE,center = TRUE)
summary(molePr)



```
<br> I scaled and centered the data to standardize it.PC1 accounts for 46% of the total variance in the data. <br>


### PCA plott and select best number of PCs for your data.

```{r }
screeplot(molePr, type = "l", npcs = 15, main = "Screeplot of the first 10 PCs")
abline(h = 1, col="blue", lty=5)
legend("topright", legend=c("Eigenvalue = 1"),
       col=c("red"), lty=5, cex=0.6)

```


```{r }
plot(molePr$x[,1],molePr$x[,2], xlab="PC1 (46.7%)", ylab = "PC2 (17.6%)", main = "PC1 / PC2 - plot")

ggplot(mole , aes(molePr$x[,1],molePr$x[,2],col=benign_malignant,fill=benign_malignant)) + 
  stat_ellipse(geom = "polygon" , col="black",alpha=0.5)+
  geom_point(shape=21,col="black")+
  xlab("Dim1 (46.7%)")+
  ylab("Dim2 (17.6%)")+
  ggtitle("2D PCA-plot from 11 feature dataset")



```

<br> I observed that the eigenvalues ​​of the first 3 components were >1. This explains a large part of the variance. So we can reduce our values from 10 to 3.It was also observed that more than 60% of the variance could be explained by the variables PC1 and PC2.That's why pc1 and pc2 components were used.The distribution of benign and malignant moles with 2 components was clearly observed. <br>











